---
title: "Motor Trend Analysis"
author: "Gazi Tarique Mahmud"
date: "12/6/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Context
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
“Quantifying how different is the MPG between automatic and manual transmissions?”
### Question
Take the mtcars data set and write up an analysis to answer their question using regression models and exploratory data analyses.

Your report must be:

Written as a PDF printout of a compiled (using knitr) R markdown document. Do not use any packages that are not in R-base or the library datasets. Brief. Roughly the equivalent of 2 pages or less for the main text. Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures. Include a first paragraph executive summary.

### Load the data

```{r}
library(datasets)
data(mtcars)
```
It consists of 32 observations on 11 variables.

[, 1] mpg Miles/(US) gallon
[, 2] cyl Number of cylinders
[, 3] disp Displacement (cu.in.)
[, 4] hp Gross horsepower
[, 5] drat Rear axle ratio
[, 6] wt Weight (lb/1000)
[, 7] qsec 1/4 mile time
[, 8] vs V/S
[, 9] am Transmission (0 = automatic, 1 = manual)
[,10] gear Number of forward gears
[,11] carb Number of carburetors

### Is an automatic or manual transmission better for MPG?
For automatic:
```{r}
summary(mtcars[mtcars$am==0,])
```
For manual:
```{r}
summary(mtcars[mtcars$am==1,])
```
Hence, the mean of mpg is greater for manual (at 24.4) than automatic (at 17.1).

### Quantify the MPG difference between automatic and manual transmissions.
```{r}
boxplot(mpg ~ am, data = mtcars, xlab = "Transmission", ylab = "Miles per gallon", main="Miles per gallon by Transmission Type")
```
Manual (represented by 1) has a higher mean for mpg than automatic (represented by 0).

### Hypothesis Testing
```{r}
aggregate(mpg~am, data = mtcars, mean)
```
The mean transmission for manual is 7.24mpg higher than automatic. Let alpha=0.5.

```{r}
auto <- mtcars[mtcars$am == 0,]
manual <- mtcars[mtcars$am == 1,]
t.test(auto$mpg, manual$mpg)
```
Since p-value = 0.001374, we reject the null hypothesis. There is a major difference between mpg of manual and automatic transmissions.

```{r}
m<-lm(mpg~am,data=mtcars)
summary(m)
```
From the above, we may conclude that automatic run at 17.15mpg, while manual have 7.24 more mpg. Also, R^2 is 0.36, hence the model only accounts for 36% variance.

Performing multivariate linear regression:
```{r}
model <- lm(mpg~am + wt + hp + cyl, data = mtcars)
anova(m,model)
```
The final model is below:
```{r}
summary(model)
```
### Conclusion
This model explains 84.9% of the variance. It may be concluded that on average, manual transmissions have 1.478 more mpg than automatic.

## APPENDIX
```{r}
plot(model)
```
Hence, the residuals are normally distributed, and homoskedastic.
